No Detectable Evidence of Genetic Associations Between Anadromous and Resident Masu Salmon Sampled within Hokkaido

### 2022-06-01 paper/paper_draft.Rmd
updated intro

---
title: "MASU Paper"
author: "William Hemstrom"
date: "11/21/2021"
output:
  word_document:
    reference_docx: "style_ref.docx"
  pdf_document:
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
    fig_width: 7
    fig_height: 6
bibliography: "Citations.bib"
csl: "council-of-science-editors-author-date.csl"
header-includes:
  - \usepackage[left]{lineno}
  - \linenumbers
  - \usepackage{setspace}\doublespacing
  - \usepackage{sectsty}\sectionfont{\fontsize{12}{12}\selectfont}
  - \usepackage{sectsty}\subsectionfont{\normalfont\itshape\fontsize{12}{12}\selectfont}
  - \usepackage[round]{natbib}
indent: true
sansfont: Times New Roman
fontsize: 12pt
---

@@

#Abstract
Within Masu salmon (Oncorhynchus masou), individuals can exhibit one of two life cycles: resident or anadromous. The expresson of either life history is mainly influenced by environmental factors such as temperature alongside other factors such as the salmon's body size, growth rate, lipid stores, and is heavily sex influenced (CITE). Here, we explore if there are genetic factors that influence the development of either residency or anadromy. In order to look for patterns of relatedness among resident and anadromous individuals, we made a PCA chart, did association tests, and made a random forest (RF). Ultimately, we did not find any significant associated genetic elements, but further research with a larger sample size is needed in order to fully assess potential genetic influences on anadromy and residency.

#Introduction

<!-- Declines in migratory fishes -->
<!-- 	Trends and potential causes. -->

<!-- Importance of Masu -->
<!-- 	Vulnerability -->
<!-- 		Very prone to fishing pressure -->
<!-- 		Very limited distribution -->

<!-- Life History -->
<!-- 	What decides if an indivividual is a resident/anadromous? -->

<!-- Genetics of resident vs anadromous -->

<!-- What did we do and find? -->
<!-- 	Thesis -->

Due to the history of hydrologic modifications such as the development of reservoirs, dams, and irrigation systems, there has been an apparent decrease in the number of migratory fishes in the recent years (CITE). In North America, for example,there have been significant decreases indiversity within pacific salmonid species due hydrologic and dam alterations (CITE), and atlantic salmon and anadromous species of sturgeons are also impacted negatively because of a loss of portions to their habitats upstream of dams (CITE). Overall, the diversity within migratory fish species have largely declined by an average of 70% between 1970 and 2016 due to these dams blocking the fishes' migratory paths (CITE- need a research paper for this number)

Masu salmon (Oncorhynchus masou, otherwise known as cherry salmon) have the most limited distribution among the Pacific salmon (CITE). Since they spend a greater portion of their life in freshwater rivers or streams than most other migratory salmon, they are particularly more exposed to fishing and sportfishing pressures (CITE).Their increased vulnerability makes it critical to understand the details that influence and control migratory life history variance in the species. Masu in particular (like Oncorhynchus mykiss, steelhead salmon/rainbow trout) are partially migratory, with large resident and anadromous populations (CITE). As is typical, anadromous masu are initially born in the rivers then migrate to the seas where they develop larger body sizes, then eventually migrate back to their natal rivers in order to breed (CITE). Resident masu in contrast sexually mature at a younger age and stay within the river systems throughout their entire lives (Morita and Nagasawa)

The decision to migrate in masu salmon is thought to be primarily influenced by environmental factors such as temperature alongside other factors such as the salmon's body size, growth rate, lipid stores, and is heavily sex influenced (CITE). Consequentially, changes in environmental factors, such as those caused by climate change, have been known to impact the decision of the masu to be either resident or anadromous (CITE). For example, X et al (YEAR) has shown that the number of resident males have increased over the last X years while the proportion of anadromous males and the proportion of delayed migrants within both females and males have decreased as the temperatures have increased, which has led to a significant decrease in the number of anadromous individuals overall.

In contrast, the genetic factors that contribute to migration in masu salmon have not been well studied.  Despite the prevalence of environmental control of migration in masu, it is not unreasonable to expect genetic elements to play some role in governing migratory decisions in the species. For example, within steelhead, there is strong evidence suggesting both an environmental and genetic basis for migratory life history. In 1944, Naeve conducted common garden and transplant experiments with sympatric rainbow trout and steelhead and found that migratory status was heritable after observing that rainbow trout offspring were considerably more likely to be resident themselves, whereas  crosses between anadromous individuals resulted in the most anadromous smolts. It is not surprising, therefore, that a recent study by Hess et al. (2016), on 15,239 single nucleotide polymorphisms (SNPs) found of three SNPs from a gene that was strongly linked with adult migration timing, and that 18 total SNPs that were connected to 60% of the trait variation. Despite this, however, growth rates are known to strongly influence anadromy in the species—individuals which grow faster as juveniles are far more likely to stay as residents than otherwise (CITE). Indeed, migration in the species seems to be controlled jointly the interplay of genetic and environmental factors:

The degree of genetic control of migration in masu, however, is still not well known. To improve our understanding of the mechanisms behind residency and anadromy in the species,  we sequenced a set of anadromous and resident masu individuals using high-resolution Restriction Associated Digest (RAD) sequencing.  However, we did not find any clear indicator of the genetic difference between ouranadromous and resident Masu salmon. While this may suggest that there is not that there is a strong genetic basis for Masu salmon migratory life history variation, further studies using a bigger sample size from a larger geographic range would need to be conducted in order to fully investigate the influences of genetics on Masu salmon migratory life cycles.

<!-- Anadromous salmon -> dams causing them to not be able to jump over dams -> is we explore the genetic influences of anadromy and residents in MASU salmon then we can tell where to focus our conservation efforts. -->

<!-- Parr size was generally larger and maturation was more common at more southern latitudes -->

<!-- Population density and temperature two most important factors affecting the growth of the fish.  -->

<!-- superior development promotes residency (early maturation) -->

<!-- need to do citations -->
```{r, include=FALSE, eval=TRUE}
dat<-read.table("../data/genotypes.geno")                        
meta<- read.table("../data/sample_meta.txt", header=TRUE)
snp.meta<- dat[,1:2]
colnames(snp.meta)<- c("chr", "position")
library(snpR)
# dat<- import.snpR.data(dat[,-c(1:2)], snp.meta = snp.meta, sample.meta = meta)
# dat<- filter_snps(dat, maf = 0.05, hwe = 0.000001)
# number_snps_after_filtering <- nsnps(dat)
# 
# #calculating genetic diversity
# dat <- calc_ho(dat, c("pop", "phen", "pop.phen"))
# get.snpR.stats(dat, c("phen", "pop","pop.phen"), stats="ho") $weighted.means
# get.snpR.stats(dat, c("pop", "phen", "pop.phen"))
# dat <- calc_fis(dat, c("pop", "phen"))
# dat <- calc_pi(dat, c("pop", "phen"))
# get.snpR.stats(dat, c("pop"), stats="pi") $weighted.means
# dat <- calc_private(dat, c("phen", "pop","pop.phen"))
# get.snpR.stats(dat, c("pop"), stats="private") $weighted.means
# get.snpR.stats(dat, c("pop", "phen", "pop.phen"))
# dat <- calc_pairwise_fst(dat, c("phen", "pop","pop.phen"), boot = 100, boot_par = 2)
# get.snpR.stats(dat, c("phen", "pop","pop.phen"), stats="fst") $weighted.means
# get.snpR.stats(dat, c("pop", "phen"), stats= c("ho", "pi", "fst")) $weighted.means
# 
# gp<- run_genomic_prediction(dat, response = "phen", iterations = 10000, burn_in = 1000, thin = 100)
# library(data.table)
# cv<- cross_validate_genomic_prediction(dat, response = "phen", iterations = 10000, burn_in = 1000, thin = 100)

@@

However, our study included samples from only a small region at the center of the masu range. It is possible that if we sampled from other areas, we could uncover loci that influence migratory life history phenotypes under environmental conditions that differ from those in our study systems. This could occur if the effect of migratory genes depends on environmental conditions (GxE). This phenomena is seen in steelhead, where variance in environmental conditions across the species’ range can determine whether life histories are more genetically or environmentally based (CITE).

>>>>

---
title: "No Detectable Evidence of Genetic Associations Between Anadromous and Resident Masu Salmon Sampled within Hokkaido"
author: "William Hemstrom"
date: "11/21/2021"
output:
  word_document:
    reference_docx: "style_ref.docx"
  pdf_document:
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
    fig_width: 7
    fig_height: 6
bibliography: "Citations.bib"
csl: "council-of-science-editors-author-date.csl"
header-includes:
  - \usepackage[left]{lineno}
  - \linenumbers
  - \usepackage{setspace}\doublespacing
  - \usepackage{sectsty}\sectionfont{\fontsize{12}{12}\selectfont}
  - \usepackage{sectsty}\subsectionfont{\normalfont\itshape\fontsize{12}{12}\selectfont}
  - \usepackage[round]{natbib}
indent: true
sansfont: Times New Roman
fontsize: 12pt
---

@@

#Abstract

Due to the history of hydrologic modifications such as the development of reservoirs, dams, and irrigation systems, there has been an apparent decrease in the number of migratory fishes in recent years (Pringle et al., 2000). In North America, for example,there have been significant decreases in diversity within pacific salmonid species due to hydrologic and dam alterations, and atlantic salmon and anadromous species of sturgeons are also impacted negatively because of a loss of portions to their habitats upstream of dams (Pringle et al., 2000). Overall, the diversity within migratory fish species have largely declined by an average of 70% between 1970 and 2016 due to these dams blocking the fishes' migratory paths (Deinet et al., 2020).

Masu salmon (Oncorhynchus masou, otherwise known as cherry salmon) have the most limited distribution among the Pacific salmon. Since they spend a greater portion of their life in freshwater rivers or streams than most other migratory salmon, they are particularly exposed to fishing and sportfishing pressures (Miyakoshi et al., 2009). Their increased vulnerability makes it critical to understand the details that influence and control migratory life history variance in the species. Masu in particular (like steelhead salmon/rainbow trout Oncorhynchus mykiss and other species to a lesser degree) are partially migratory, with large resident and anadromous populations (Morita et al., 2014). As is typical, anadromous masu are initially born in the rivers then migrate to the seas where they develop larger body sizes, then eventually migrate back to their natal rivers in order to breed (Dodson et al., 2013). Resident masu in contrast sexually mature at a younger age and stay within the river systems throughout their entire lives (Morita and Nagasawa).

The decision to migrate in masu salmon is thought to be primarily influenced by environmental factors such as temperature alongside other factors such as the salmon's body size, growth rate, lipid stores, and is heavily sex influenced (Morita et al., 2014; Arostegui & Quinn, 2019). Consequently, changes in environmental factors, such as those caused by climate change, have been known to impact the decision of the masu to be either resident or anadromous (Morita et al., 2014). For example, Morita et al (2014) has shown that the number of resident males have increased while the proportion of anadromous males and the proportion of delayed migrants within both females and males have decreased as the temperatures have increased, which has led to a significant decrease in the number of anadromous individuals overall.

In contrast, the genetic factors that contribute to migration in masu salmon have not been well studied.  Despite the prevalence of environmental control of migration in masu, it is not unreasonable to expect genetic elements to play some role in governing migratory decisions in the species. For example, within steelhead, there is strong evidence suggesting both an environmental and genetic basis for migratory life history. In 1944, Naeve conducted common garden and transplant experiments with sympatric rainbow trout and steelhead and found that migratory status was heritable after observing that rainbow trout offspring were considerably more likely to be resident themselves, whereas  crosses between anadromous individuals resulted in the most anadromous smolts. It is not surprising, therefore, that a recent study by Hess et al. (2016), on 15,239 single nucleotide polymorphisms (SNPs) found of three SNPs from a gene that was strongly linked with adult migration timing, and that 18 total SNPs that were connected to 60% of the trait variation. Despite this, however, growth rates and body condition are known to influence anadromy in the species—individuals which grow faster as juveniles are more likely to stay as residents than otherwise (Kendall et al., 2014). Growth rates are also known to be influenced by large genetic elements that differ between anadromous and resident individuals, implying that genetics may contribute to migration indirectly as well (Miller et al., 2011). Taken together, it seems clear that migration in steelhead is  jointly controlled by the interplay of genetic and environmental factors.

The degree of genetic control of migration in masu, however, is still not well known. To improve our understanding of the mechanisms behind residency and anadromy in the species, we sequenced a set of anadromous and resident masu individuals using high-resolution Restriction Associated Digest (RAD) sequencing.  However, we did not find any clear indicator of the genetic difference between our anadromous and resident masu salmon. While this may suggest that there is not that there is a strong genetic basis for masu salmon migratory life history variation, further studies using a bigger sample size from a larger geographic range would need to be conducted in order to fully investigate the influences of genetics on Masu salmon migratory life cycles.

```{r, include=FALSE, eval=TRUE}
dat<-read.table("../data/genotypes.geno")                        
meta<- read.table("../data/sample_meta.txt", header=TRUE)
snp.meta<- dat[,1:2]
colnames(snp.meta)<- c("chr", "position")
library(snpR)
# dat<- import.snpR.data(dat[,-c(1:2)], snp.meta = snp.meta, sample.meta = meta)
# dat<- filter_snps(dat, maf = 0.05, hwe = 0.000001)
# number_snps_after_filtering <- nsnps(dat)
# 
# #calculating genetic diversity
# dat <- calc_ho(dat, c("pop", "phen", "pop.phen"))
# get.snpR.stats(dat, c("phen", "pop","pop.phen"), stats="ho") $weighted.means
# get.snpR.stats(dat, c("pop", "phen", "pop.phen"))
# dat <- calc_fis(dat, c("pop", "phen"))
# dat <- calc_pi(dat, c("pop", "phen"))
# get.snpR.stats(dat, c("pop"), stats="pi") $weighted.means
# dat <- calc_private(dat, c("phen", "pop","pop.phen"))
# get.snpR.stats(dat, c("pop"), stats="private") $weighted.means
# get.snpR.stats(dat, c("pop", "phen", "pop.phen"))
# dat <- calc_pairwise_fst(dat, c("phen", "pop","pop.phen"), boot = 100, boot_par = 2)
# get.snpR.stats(dat, c("phen", "pop","pop.phen"), stats="fst") $weighted.means
# get.snpR.stats(dat, c("pop", "phen"), stats= c("ho", "pi", "fst")) $weighted.means
# 
# gp<- run_genomic_prediction(dat, response = "phen", iterations = 10000, burn_in = 1000, thin = 100)
# library(data.table)
# cv<- cross_validate_genomic_prediction(dat, response = "phen", iterations = 10000, burn_in = 1000, thin = 100)

@@

However, our study included samples from only a small region at the center of the masu range. It is possible that if we sampled from other areas, we could uncover loci that influence migratory life history phenotypes under environmental conditions that differ from those in our study systems. This could occur if the effect of migratory genes depends on environmental conditions (GxE). This phenomena is seen in steelhead, where variance in environmental conditions across the species’ range can determine whether life histories are more genetically or environmentally based (Kendall et al., 2014).
